---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mlsubtyping

Provide functionality for cancer subtyping using existing published methods or machine learning based on TCGA data.

Currently support mRNA subtyping:

  * LUSC using [nearest centroids method](https://clincancerres.aacrjournals.org/content/16/19/4864) or random forest method by training TCGA data.
  * LUAD using [nearest centroids method](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0036530) or random forest method by training TCGA data.
  * HNSC using [nearest centroids method](https://www.nature.com/articles/nature14129) or random forest method by training TCGA data.
  * BLCA using random forest (rf) method by training TCGA data.
  * STAD using random forest (rf) method by training TCGA data.
  * ESCA using random forest (rf) method by training TCGA data.
  * ESCC using random forest (rf) method by training TCGA data.
  * BRCA using PAM50 method based on R package genefu.
  
## Latest release

0.2.4
<!-- badges: start -->
<!-- badges: end -->

## Installation

You can install the released version through:

``` r
install.packages("mlsubtyping_0.2.4.tar.gz", repos = NULL, type="source")
```
Or

``` r
R CMD INSTALL mlsubtyping_0.2.4.tar.gz
```
## Example

This is a basic example for predicting the subtypes for (Lung Squamous Cell Carcinoma) LUSC.

### Predict LUSC mRNA Expression Subtypes using [nearest centroids method](https://clincancerres.aacrjournals.org/content/16/19/4864)

```{r wilkerson, message=FALSE}
library(mlsubtyping)
library(tidyverse)
data <- get_median_centered(example_fpkm)
data <- assays(data)$centered
rownames(data) <- rowData(example_fpkm)$external_gene_name
# use default wilkerson's method
output1 <- CentroidsSubtype(data, disease = 'LUSC')
table(output1@subtypes)
```

### Using random forest model by training TCGA LUSC data
```{r rf, message=FALSE}
output2 <- MLSubtype(data, disease = 'LUSC', method = 'rf', seed = 123)
table(output2@subtypes)
```

### Check the consistance between two methods
```{r confusion, message=FALSE}
confusionMatrix(as.factor(tolower(output1@subtypes)),
                as.factor(tolower(output2@subtypes)))
```
### Plot important genes
```{r varImp, message=FALSE}
vi <- varImp(output2@method, scale = TRUE)
plot(vi, top = 20) 
```

### Plotheat map
```{r heatmap, message=FALSE}
PlotHeat(object = output2, set = 'both', fontsize = 10,
        show_rownames = FALSE, show_colnames = FALSE)
```

## Report bugs or issues at [here](https://dgit.eacc.ds.quintiles.com/u1072932/mlsubtyping/issues)
